---
title: "STAT S460F final project"
author: "Siu Kwok Lai, 12354991"
date: "2020/11/27"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# List of packages
```{r packages, eval=FALSE}
install.packages("tidyr")
install.packages("devtools")
install_github("vqv/ggbiplot")
install.packages("ISLR")
install.packages("ggplot2")
install.packages("epiDisplay")
install.packages("leaps")
install.packages("PerformanceAnalytics")
install.packages("factoextra")
install.packages("ggbiplot")
install.packages("ggpubr")
install.packages("dplyr")
install.packages("ISLR")
install.packages("e1071")
install.packages("car")
install.packages("PreProcess")
install.packages("glmnet")
install.packages("pls")
```

# List of libraries 
```{r acces lib, message=FALSE}
library(tidyr) ## Handle data
library(devtools)## Important to install ggbiplot
library(ISLR) ## Data Mining Package
library(ggplot2) ## Graphs
library(epiDisplay) ## One way Analysis
library(leaps) ### Predict
library(PerformanceAnalytics) ## Correlation
library(factoextra) ## Extract and Visualize result
library(ggbiplot) ## Graphs
library(ggpubr) ## qqplot
library(dplyr) ## Manipulation
library(ISLR) ## Data
library(e1071) ## SVM
library(car) ## Simple analysis
library(PreProcess) ## Handle data
library(glmnet) ## Graphs
library(pls) ## PLS, PLSR
```

# Data Fetching
```{r fetching}
mydata = data.frame(read.csv("C:\\Users\\LSK\\OneDrive - The Open Unviersity of Hong Kong\\r project\\Pizza.csv"))
```

# 4. Data description

## Dataset(Mydata) Manipulation
```{r dataset manipulation}
dim(mydata)
sum(is.na(mydata)) ## count if missing, and no missing data existing
head(mydata)
str(mydata) 

names(mydata)[3:9] = c("water", "protein", "fat", "ash", "sodium", "carbohydrates", "calories")

mydata$brand = as.factor(mydata$brand)
mydata$id = as.integer(mydata$id)
str(mydata)

mydata= mydata[,-2]
str(mydata)
```


## Split data
```{r split mydata}
set.seed(1)
train = mydata %>%
sample_frac(0.5)
test = mydata %>%
setdiff(train)

dim(train)
dim(test)
```

# 5. Exploratory and descriptive analysis of data

## Descriptive and exploratory analysis of each variable 
```{r Descriptive1}

tab1(mydata$brand, cum.percent = TRUE,main = "brand")

summary(mydata$calories)
hist(mydata$calories,main="Histogram of calories (per 100 grams)", freq=F, breaks=10, ylim=c(0,1))
lines(density(mydata$calories), col="red")
lines(seq(1, 5, by=.1), dnorm(seq(1, 5, by=.1),
      mean(mydata$calories), sd(mydata$calories)), col="blue")

summary(mydata$water)
hist(mydata$water,main="Histogram of water (per 100 grams)", freq=F, breaks=10, xlim = c(20,60))
lines(density(mydata$water), col="red")
lines(seq(20, 60, by=.5), dnorm(seq(20, 60, by=.5),
      mean(mydata$water), sd(mydata$water)), col="blue")

summary(mydata$protein)
hist(mydata$protein,main="Histogram of protein (per 100 grams)", freq=F, xlim = c(1,30))
lines(density(mydata$protein), col="red")
lines(seq(0, 30, by=.1), dnorm(seq(0, 30, by=.1),
      mean(mydata$protein), sd(mydata$protein)), col="blue")

summary(mydata$fat)
hist(mydata$fat,main="Histogram of fat (per 100 grams)", freq=F, ylim=c(0,0.1))
lines(density(mydata$fat), col="red")
lines(seq(0, 50, by=.5), dnorm(seq(0, 50, by=.5),
      mean(mydata$fat), sd(mydata$fat)), col="blue")

summary(mydata$ash)
hist(mydata$ash,main="Histogram of ash (per 100 grams)", freq=F, xlim = c(0,6))
lines(density(mydata$ash), col="red")
lines(seq(0, 6, by=.1), dnorm(seq(0, 6, by=.1),
      mean(mydata$ash), sd(mydata$ash)), col="blue")

summary(mydata$sodium)
hist(mydata$sodium,main="Histogram of sodium (per 100 grams)", freq=F,xlim = c(0,2), ylim=c(0,3.5))
lines(density(mydata$sodium), col="red")
lines(seq(0, 2, by=.1), dnorm(seq(0, 2, by=.1),
      mean(mydata$sodium), sd(mydata$sodium)), col="blue")

summary(mydata$carbohydrates)
hist(mydata$carbohydrates,main="Histogram of carbohydrates (per 100 grams)", freq=F)
lines(density(mydata$carbohydrates), col="red")
lines(seq(0, 50, by=.5), dnorm(seq(0, 50, by=.5),
      mean(mydata$carbohydrates), sd(mydata$carbohydrates)), col="blue")


boxplot(mydata)

chart.Correlation(mydata[, -1], histogram=TRUE, pch=19)

```


## Diagnostic plots - multiple linear regression model

```{r lm}
data.lm = lm(calories ~ ., data = mydata)
plot(data.lm) 

```

# 6. Analysis of data based on given objectives

## Objective 1

### Ridge/Lasso preparation 
```{r R and L}

x_train = model.matrix(calories~., train[,-1])[,-1]
x_test = model.matrix(calories~., test[,-1])[,-1]

y_train = train[-1] %>%
select(calories) %>%
unlist() %>%
as.numeric()

y_test = test[-1] %>%
select(calories) %>%
unlist() %>%
as.numeric()


grid = 10^seq(10, -2, length = 100)
```

### Ridge regression model 
```{r ridge}
ridge_mod = glmnet(x_train, y_train, alpha=0, lambda = grid)
ridge_pred = predict(ridge_mod, s = 4, newx = x_test)

mean((ridge_pred - y_test)^2)
plot(ridge_mod)

set.seed(1)
cv.out = cv.glmnet(x_train, y_train, alpha = 0) # Fit ridge regression model on training 
bestlam = cv.out$lambda.min # Select lambda that minimizes training MSE
bestlam
plot(cv.out) # Draw plot of training MSE as a function of lambda

ridge_mod=glmnet(x_train, y_train, alpha = 0, lambda = bestlam)
coef(ridge_mod)

ridge_pred = predict(ridge_mod, s = bestlam, newx = x_test) # Use best lambda to predict 
mean((ridge_pred - y_test)^2) # Calculate test MSE


```

### Lasso regression model 
```{r Lasso}
lasso_mod = glmnet(x_train, y_train, alpha=1, lambda = grid)
lasso_pred = predict(lasso_mod, s=4, newx = x_test)

mean((lasso_pred - y_test)^2)
plot(lasso_mod)

set.seed(1)
cv.out = cv.glmnet(x_train, y_train, alpha=1)
bestlam = cv.out$lambda.min
bestlam
plot(cv.out)

lasso_mod=glmnet(x_train, y_train, alpha = 1, lambda = bestlam)
coef(lasso_mod)

lasso_pred = predict(lasso_mod, s = bestlam, newx = x_test)
mean((lasso_pred - y_test)^2) # MSE

lasso_lm = lm(calories~water+protein+fat+ash, data=test)
summary(lasso_lm)
vif(lasso_lm)

```

### PCR
```{r PCR}

set.seed(1)
pcr_fit = pcr(calories~., data = train[,-1], scale = TRUE, validation = "CV")
summary(pcr_fit)


validationplot(pcr_fit, val.type = "MSEP")
plot(pcr_fit, "loadings") 
plot(pcr_fit, "coef") # Coefficients

pcr_pred = predict(pcr_fit, x_test, ncomp=2)
mean((pcr_pred-y_test)^2)
coef(pcr_fit,2)


```

### PLSR
```{r PLSR}
set.seed(1)
pls_fit = plsr(calories~., data = train[,-1], scale = TRUE, validation = "CV")
summary(pls_fit)

validationplot(pls_fit, val.type = "MSEP")
pls_pred = predict(pls_fit, x_test, ncomp = 2)
mean((pls_pred - y_test)^2) # MSE
coef(pls_fit,2)

```



## Objective 2

### PCA
```{r pca}
train.pca = prcomp(train[,c(-1,-8)], center = TRUE,scale. = TRUE)
summary(train.pca)
summary(train.pca)$rotation
##summary(mydata.pca)$x
fviz_eig(train.pca)

ggbiplot(train.pca, ellipse=TRUE, groups=train$brand)

fviz_pca_var(train.pca,
col.var = "contrib", # Color by contributions to the PC
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE # Avoid text overlapping
)

fviz_pca_biplot(train.pca, repel = TRUE,
col.var = "#2E9FDF", # Variables color
col.ind = "#696969" # Individuals color
)

# Eigenvalues
eig.val= get_eigenvalue(train.pca)
eig.val

# Results for Variables
res.var= get_pca_var(train.pca)
res.var$coord # Coordinates

res.var$contrib # Contributions to the PCs
res.var$cos2 # Quality of representation

```

### SVM
```{r svm}
traning_brand = train$brand
training_data = train[,c(4,7)]
training_data = data.frame(training_data, brand = as.factor(traning_brand))

testing_brand = test$brand
testing_data = test[,c(4,7)]
testing_data = data.frame(testing_data, brand = as.factor(testing_brand))


ggplot(data.frame(training_data), aes(training_data$carbohydrates, training_data$fat, colour = traning_brand)) +
geom_point()


```

### 10-folds cross-validation to find best cost and gamma
```{r ten}
brand = training_data$brand
set.seed(1)
tune_out=tune(svm,
              factor(brand)~.,
              data=training_data,
              kernel="radial",
              ranges = 
                list(cost = c(0.1,1,10,50,100,500,1000),
                     gamma = c(0.5,1,2,3,4,5,6,7,8,9,10)))
              

bestmod=tune_out$best.model
summary(bestmod)
plot(bestmod,training_data)


```

### Best values on testing dataset
```{r SVM}
svmfit = svm(testing_data$brand~., data = testing_data, kernel = "radial", cost = 50, gamma = 10)
plot(svmfit, testing_data)

```

### The percentage of accuracy
```{r final}
pred_table = table(true = testing_data$brand, pred = predict(tune_out$best.model, newdata = testing_data))
pred_table
accuracy = sum(diag(pred_table))/nrow(testing_data)
accuracy
error = 1 - accuracy
error

```

### bar chart
```{r bar}

new_fat = aggregate(mydata$fat, by=list(brand=mydata$brand), FUN=sum)
new_carbo = aggregate(mydata$carbohydrates, by=list(brand=mydata$brand), FUN=sum)

new_plotdata = cbind(new_fat,new_carbo[,-1])
names(new_plotdata)[c(2,3)] = c('fat','carbohydrates')
new_plotdata
t(new_plotdata)

dat = data.frame(
  fat = new_plotdata$fat,
  carbohydrates = new_plotdata$carbohydrates,
  brands = as.factor(new_plotdata$brand)
)

dat_long <- dat %>%
  gather("nutrients", "grams", -brands)

ggplot(dat_long, aes(x = brands, y = grams, fill = nutrients)) + ggtitle("Total grams of effective nutrients by brands") +
  geom_col(position = "dodge")

```

